---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggtime

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ggtime)](https://CRAN.R-project.org/package=ggtime)
[![R-CMD-check](https://github.com/mitchelloharawild/ggtime/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mitchelloharawild/ggtime/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The ggtime package extends the capabilities of 'ggplot2' by providing 
grammatical elements and plot helpers designed for visualizing time series 
patterns. These functions use calendar structures implemented in the mixtime 
package to help explore common time series patterns including trend, 
seasonality, cycles, and holidays.

The plot helper functions make use of the tsibble data format in order to 
quickly and easily produce common time series plots. These plots can also be
constructed with the underlying grammar elements, which allows greater 
flexibility in producing custom time series visualisations.

## Installation

You can install the **stable** version from
[CRAN](https://cran.r-project.org/package=ggtime):

``` r
install.packages("ggtime")
```

You can install the development version of ggtime from [GitHub](https://github.com/) with:

```r
# install.packages("remotes")
remotes::install_github("tidyverts/ggtime")
```

## Example

The simplest time series visualisation is the time plot, which can be created with the `autoplot()` plot helper function on a tsibble object.

```{r timeplot}
#| warning: false
#| message: false
library(ggtime)
library(tsibble)
library(ggplot2)
tsibbledata::aus_production %>% 
  autoplot(Bricks)
```

To view the shape of the annual seasonal pattern, a seasonal plot created with `gg_season()` is commonly used. This makes it easier to identify the peaks, troughts, and overall shape of the seasonality.

```{r seasonplot}
#| warning: false
tsibbledata::aus_production %>% 
  gg_season(Bricks)
```


```{r}
#| include: false
#| eval: false
cal_trans_x <- function() {
  scales::trans_new(
    name = "calendar",
    transform = ggtime:::calendar_wrap,
    inverse = identity,
    breaks = scales::breaks_pretty(),
    domain = c(0, 60*60*24*7)
  )
}

pedestrian[with(pedestrian, Sensor == "Southern Cross Station" & Date < "2015-03-01"),] |> 
  autoplot(Count) +
  # coord_calendar(xlim = c(Sys.time(), Sys.Date() + lubridate::days(1)))
  ggplot2::coord_trans(x = cal_trans_x(), xlim = as.POSIXct(c("2024-03-25 00:00:00", "2024-03-31 23:59:59"))) +
  scale_x_datetime(date_breaks = "day", date_labels = "%a")

pedestrian[with(pedestrian, Sensor == "Southern Cross Station" & Date < "2015-03-01"),] |> 
  ggplot(aes(x = Date_Time, y = Count)) + 
  geom_path() +
  ggplot2::scale_x_continuous(transform = cal_trans_x(), limits = as.POSIXct(c("2024-03-25 00:00:00", "2024-03-31 23:59:59"))) + 
  coord_polar()
```

